---
title: "Example2_MetaAnalysis"
author: "Kalani Ratnasiri"
date: "2025-02-24"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multi-dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=8, fig.height=6, warning = FALSE) 
```

## Install packages as necessary
```{r installs, include = FALSE}
# install.packages("https://seurat.nygenome.org/src/contrib/ifnb.SeuratData_3.0.0.tar.gz", repos = NULL, type = "source")
```

```{r setup, message = FALSE}
library(scMetaIntegrator)
library(Seurat)
library(ifnb.SeuratData)
library(MetaIntegrator)
library(multtest)
library(magrittr)
library(future.apply)
```

## Step 0a: Data Loading + Pre-processing
To get data to work with, we are going to take a publicly available dataset through SeuratData. We will assign cells to 4 pseudo donors to have a variable to pair on. 
```{r, message=FALSE}
data("ifnb")
ifnb<-UpdateSeuratObject(ifnb)
ifnb = NormalizeData(ifnb)
ifnb@meta.data$pseudoDonor = rep(c(1,2,3,4), nrow(ifnb@meta.data)/2)[c(1:nrow(ifnb@meta.data))] ## randomly assign donor
ifnb@meta.data$pseudoDonor = paste0("pseudoDonor_",ifnb@meta.data$pseudoDonor)
ifnb@meta.data$sample = paste0(ifnb@meta.data$pseudoDonor,"_",ifnb@meta.data$stim)
```

## Step 0b: Test sample removal
We will remove a sample to demonstrate how to handle sample imbalance
```{r}
ifnb = subset(ifnb, subset = sample != "pseudoDonor_4_STIM")
table(ifnb$pseudoDonor, ifnb$stim)
```


## Step 1: Confirm pairing structure
Next, we want to make sure that we have complete pairing in this dataset. 
```{r}
confirmPairs(ifnb, uniqueSampleID = "sample", pairingColumn = "pseudoDonor")
```

## Step 2: Remove unpaired samples
Based on step 1, we have incomplete pairing. Therefore, we're going to subset out object to only keep complete pairs
```{r}
ifnb_complete = completePairSubset(ifnb, uniqueSampleID = "sample", pairingColumn = "pseudoDonor")
table(ifnb$pseudoDonor, ifnb$stim)
```

## Step 3: Run thresholding for genes that are expressed across 1% of cells in each pairing. 
```{r}
thresholdedGenes = thresholdGenesAcrossPairs(
  seuratObject = ifnb_complete,
  uniqueSampleID = "sample",
  pairingColumn = "pseudoDonor",
  percentThreshold = 0.01
)
length(thresholdedGenes)
```

## Step 4: Run scMI and generate the metaObject object 
```{r}
scMIObj = generateFullMA(
  seuratObject = ifnb_complete,
  uniqueSampleID = "sample",
  pairingColumn = "pseudoDonor",
  comparisonGroupColumn = "stim",
  groupOne = "STIM",
  groupTwo = "CTRL",
  thresholdGenes = thresholdedGenes
)
```

## Step 5: Get detailed DE report
```{r}
scMI_output = scMIObj$metaAnalysis$pooledResults
```

## Step 6: Below are ways to analyze the data using the MetaIntegrator package
At this moment, all MetaIntegrator functions can be performed on the metaObject generated in Step 4. A look at available methods that can be performed on this MetaIntegrator object can be found in the [MetaIntegrator vignette](http://cran.nexr.com/web/packages/MetaIntegrator/vignettes/MetaIntegrator.html). 

### Step 6a: Forest plot of top DEG gene
```{r}
topDEG = scMI_output[scMI_output$effectSizeFDR == 0 & scMI_output$tauSquared == 0,]
topDEG = topDEG[order(topDEG$effectSize, decreasing = TRUE),]$gene[1]
forestPlot(scMIObj, topDEG)
```

### Step 6b: Generate gene score for IFNB per-cell using genes with FDR < 0.0001 & save score for each cell
```{r}
scMIObj = filterGenes(scMIObj, isLeaveOneOut = TRUE, FDRThresh = .0001)
scMIObj$originalData$pseudoDonor_1$pheno$score = calculateScore(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_1)
scMIObj$originalData$pseudoDonor_2$pheno$score = calculateScore(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_2)
scMIObj$originalData$pseudoDonor_3$pheno$score = calculateScore(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_3)
```

## Step 6c: Generate violin plots for identified gene signature per-sample sample
```{r}
violinPlot(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_1, labelColumn = 'stim')
violinPlot(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_2, labelColumn = 'stim')
violinPlot(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_3, labelColumn = 'stim')
```

## Step 6d: Generate rocPlot for one sample using gene signature
```{r}
rocPlot(scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`, scMIObj$originalData$pseudoDonor_1, title = "ROC plot for discovery psudoDonor_1, FDR: 0.0001")
```

## Step 6e: Generate pooled ROC across all pseudoDonors using gene signature
```{r}
pooledROCPlot(scMIObj, scMIObj$filterResults$`FDR1e-04_es0_nStudies1_looaTRUE_hetero0`)
```


##
