---
title: "Example1_MetaAnalysis"
author: "Kalani Ratnasiri"
date: "2025-02-24"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example1_MetaAnalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=8, fig.height=6, warning = FALSE) 
```

## Install packages as necessary
```{r installs, include = FALSE}
# install.packages("https://seurat.nygenome.org/src/contrib/ifnb.SeuratData_3.0.0.tar.gz", repos = NULL, type = "source")
```

```{r setup, message = FALSE}
library(scMetaIntegrator)
library(Seurat)
library(ifnb.SeuratData)
library(MetaIntegrator)
library(multtest)
library(magrittr)
library(future.apply)
## packages for extra visualization steps
library(ggplot2)
library(EnhancedVolcano)
```

## Step 0: Data Loading + Pre-processing
To get data to work with, we are going to take a publicly available dataset through SeuratData. We will assign cells to 3 pseudo donors to have a variable to pair on. 
```{r, message=FALSE}
data("ifnb")
ifnb<-UpdateSeuratObject(ifnb)
ifnb = NormalizeData(ifnb)
ifnb@meta.data$pseudoDonor = rep(c(1,2,3), nrow(ifnb@meta.data)/2)[c(1:nrow(ifnb@meta.data))] ## randomly assign donor
ifnb@meta.data$pseudoDonor = paste0("pseudoDonor_",ifnb@meta.data$pseudoDonor)
ifnb@meta.data$sample = paste0(ifnb@meta.data$pseudoDonor,"_",ifnb@meta.data$stim)
```


## Step 1: Confirm pairing structure
Next, we want to make sure that we have complete pairing in this dataset. 
```{r}
confirmPairs(ifnb, uniqueSampleID = "sample", pairingColumn = "pseudoDonor")
```

## Step 2: Run thresholding for genes that are expressed across 1% of cells in each pairing. 
```{r}
thresholdedGenes = thresholdGenesAcrossPairs(
  seuratObject = ifnb,
  uniqueSampleID = "sample",
  pairingColumn = "pseudoDonor",
  percentThreshold = 0.01
)
length(thresholdedGenes)
```


## Step 3: Now we want to run scMetaIntegrator and get the simplest DE output.
### For calculating effect size changes, groupOne is the group we're defining changes distinguishing these cells from groupTwo (usually the control sample)
```{r runscMI}
scMI_output = runscMI(
  seuratObject = ifnb,
  uniqueSampleID = "sample",
  pairingColumn = "pseudoDonor",
  comparisonGroupColumn = "stim",
  groupOne = "STIM",
  groupTwo = "CTRL",
  thresholdGenes = thresholdedGenes
)
```

## Step 4: Below are basic checks & visualizations of the output data
### 4a: Get the number of significant DEGs
```{r}
table(scMI_output$effectSizeFDR < 0.05)
```


### 4b: Get a simple barplot of DEGs
```{r}
scMI_output$direction = ifelse(scMI_output$effectSize > 0, "positive","negative")
scMI_output$direction = ifelse(scMI_output$effectSize == 0, "0",scMI_output$direction)
scMI_output$significant = ifelse(scMI_output$effectSizeFDR < 0.05, "TRUE","FALSE")

ggplot(scMI_output, aes(x=significant, fill = direction)) +
  geom_bar() + theme_bw() + xlab("FDR < 0.05") + ylab("# of DEGs") +
  labs(fill="sign(effectSize)") + scale_fill_manual(values = c("#16317DFF", "#A40000FF"))
```


### 4b: Get a simple volcano plot of DEGs
```{r}
EnhancedVolcano(scMI_output,
    lab = scMI_output$gene,
    x = 'effectSize',
    y = 'effectSizeFDR',
    pCutoff = 10e-32,
    FCcutoff = 0.5,
    pointSize = 3.0,
    labSize = 6.0)
```
